Refinements for open automata

### 2022-07-17 2021-Refinement/paper/refinementarticle.tex
related work

@@

The bisimulation relation that we consider is the one that makes $A_1$ and $A_2$ bisimilar, complemented with identity of configurations for $A_3$.
The proof relies on the fact that, by construction, all transitions of the composed automaton $A_1\subst{A_3}{k}$ are specified by open transitions of $A_1$. For the transitions that do not involve hole $k$, the transition of $A_1\subst{A_3}{k}$ is the same and simulation between $A_1$ and $A_2$ allows us to conclude directly. If the hole $k$ is involved the considered relation  implies that valuations in $A_3$ are equal (i.e., the value for each variable are the same in both valuations), after a transition we should obtain ``equal'' valuations because post-conditions are deterministic.
\begin{example} The composition of the refined \texttt{database} and the \texttt{counter} is again an open automata depicted in Fig.~\ref{Fig:ComposeRefine}. As stated by the theorem, this automaton is a refinement of the automaton of the figure Fig.~\ref{Def:CompOA}, the automaton resulting from the composition of the abstract \texttt{database}.
\end{example}

@@

\TODO{[RABEA] RW}

@@

Previous work on open automata focused on equivalence relations compatible with composition.
In an article by Hou, Zechen and Madelaine \cite{10.1145/3372884.3373161}, a computable bisimulation is introduced and proved equivalent to the previous bisimulation already introduced.
In a more recent work by Ameur-Boulifa, Henrio and Madelaine \cite{fhbisim}, a weak version of the bisimulation on open automata is introduced.
These works differ from ours because the relation introduced in this report is a refinement relation in the form of a simulation and not a bisimulation.
Also we do not have results as strong as computability neither a weak version able to tackle silent actions.

\TODO{Some related work on other models than open automata introduce refinement relations.
In a chapter by Bellegarde, Julliand and Kouchnarenko \cite{10.1007/3-540-46428-X_19}, a simulation relation on transition systems is introduced.
This simulation encompass action refinement, is able to deal with silent actions and is compatible with parallel composition.
Here the refinement relation does not consider action refinement as valid but it should be done in future work.
Also they check how LTL properties are preserved or combined using their refinement which we do not do.
However their model is less expressive: the transition system model is less expressive than open automata and the parallel composition is less expressive than composition on open automata.
In a later report by Kouchnarenko and Lanoix \cite{10.1007/978-3-540-70881-0_26}, the refinement relation they introduce is on LTS (labelled transition systems).
Their relation additionally prevents deadlock and livelocks.
The composition is also extended to synchronised composition which is more expressive.
In our work we also deal with deadlocks but not with livelocks since the latter arise only with silent actions.
This work is closer than the previous one to what we do here, still open automata are more expressive than LTS and composition is more general than synchronised composition.}

A refinement relation on models nearer to open automata is introduced in an article by Zhang, Meng and Lo \cite{Zhang2014}.
In their article they work with transition systems with variable which makes the state space potentially infinite.
This aspect is also present in open automata.
They show how invariants are composed.
By relation on these invariants they introduce several refinement relations.
We could have done something similar for non-blocking composition reachability predicates which are introduced in Section \ref{sec:proofelts}.

On the deadlock prevention aspect, an article by Dihego, Sampaio and Oliveira \cite{DIHEGO2020110598} present a refinement relation on process algebra (translated to LTS).
This refinement relation is a special case of inheritance and prevents the introduction of deadlocks.
Their refinement and inheritance are quite the opposite of our refinement in terms of new behaviours.
They have channels, interfaces, inputs and outputs, which in the open automata model can be compared to action labels, holes and action data for both inputs and outputs.
They have a rich composition as open automata but the introduction of deadlock is already prevented by a well chosen set of composing operations.
Also their composition is slightly different than the one on open automata because they can cause loops by linking two channels of the same process, where in open automata the composition makes an oriented tree.
In their model there is an explicit deadlock and a successful termination where in open automata there are no explicit termination.
We define a deadlock as a configuration without possible transition and assume what is a deadlock when comparing the open automata.
To define their refinement and inheritance relation they use trace and failure semantics, which are weaker than (bi)simulations \cite{10.5555/640428.640430} and could break with open automata composition.

@@

\newpage
\TODO{[RABEA] proof}
Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. According to the refinement relation between  $A_1$ and $A_2$ we have for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}	
According to the refinement  relation between   $A_2$ and $A_3$, we know the following is satisfied, in particular for  the set $(t_{2pxy})^{x\in X_p}$:
\begin{multline*}
\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_{2px}, t_3} = \mpar{\OTx{2px}{}{2px}{2px}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2pxy} = \OTx{2px}{y}{2pxy}{2pxy} \in \fOT{s_{2p}}}^{y \in Y},
V_2 \uplus V_3\uplus \biguplus_{y\in Y}\fvars{t_{2pxy}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{y\in Y}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}\\
\end{multline*}
Yet according to the refinement  relation between  $A_1$ and $A_2$ the following is satisfied, in particular over the family of transitions in $Y$:
\begin{multline*}
\bigforall \mpar{s_1, s_{2p}} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_{2pxy}} = \mpar{\OTx{1}{}{1}{1}, \OTx{2pxy}{}{2pxy}{2pxy}} \in \fOT{s_1} \times \fOT{s_{2p}} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1z} = \OTx{1}{z}{1z}{1z} \in \fOT{s_1}}^{z \in Z},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pxyj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}
\end{multline*}

From previous formulas we get:

\begin{multline*}
\mpar{R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy}} \wedge \mpar{ R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 } 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}} \wedge \\\bigsymb{\bigwedge_{y\in Y}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}
\end{multline*}

--------------------------------------------------

Yet according to the refinement  relation between   $A_2$ and $A_3$ the following is satisfied:
\begin{multline*}
\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_{2p}, t_3} = \mpar{\OTx{2p}{}{2p}{2p}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2px} = \OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{y \in Y},
V_2 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{2px}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_2, s_3} \wedge 	g_{2p}\wedge g_3 \wedge \alpha_2 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_{2p}=\alpha_{2p x}  \land J'_{2px}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2xj}\land g_{2p x}\land
 R_{23}\mpar{s'_{2p x}, s'_3}\psubst{\psi_{2p x} \uplus \psi_3}}\\
\end{multline*}
For any open transition $t_{2px}$ originating from $s_{2p}$ we have 
the following by definition of refinement between  $A_2$ and $A_3$:

\begin{align*}
 \OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}} 
\end{align*}
there exists an indexed family of OTs originating from $s_3$: 
\begin{align*}
\mpar{\OTx{3}{pxy}{3pxy}{3pxy} \in \fOT{s_3}}^{y \in Y} 
\end{align*}
such that $\forall y \in Y, J'_{2px} \cap H = J'_{3py} \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_{2px}} \vdash\\
R_{23}\mpar{s_{2p}, s_{3}} \wedge g_{2px} \implies  \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H} \beta_{2pxj}=\beta_{3pxyj} \nwedge \\
			g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}  
\end{multline*}

Based on             the above two properties, we get:
\begin{multline*}
R_{12}\mpar{s_1, s_{2p}} \wedge  R_{23}\mpar{s_{2p}, s_{3}} \wedge	g_1 \wedge g_{2p} \wedge \alpha_1 = \alpha_{2p} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2p}=\alpha_{3py} \wedge \bigwedge_\subbox{j \in J'_{3py} \cap H} \beta_{2pj}=\beta_{3pyj} \nwedge \\
			g_{3py} \wedge R_{23}\mpar{s'_{2p}, s'_{3y}}\psubst{\psi_{2p} \uplus \psi_{3py}}
		\end{array}}\bigwedge
\mpar{\everymath{\displaystyle}\begin{array}{l}		
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_{2p} \land \bigwedge_\subbox{j \in J'_{2p}} \beta_{2pj} = \beta_{1xj} \land\\ g_{1 x} \land
 R_{12}\mpar{s'_{1 x}, s'_{2p}}\psubst{\psi_{1 x} \uplus \psi_{2p}}
 \end{array}}
 }
\end{multline*}
By  rearranging the terms in the formula we get:
\begin{multline*}
R_{12}\mpar{s_1, s_{2p}} \wedge  R_{23}\mpar{s_{2p}, s_{3}} \wedge	g_1 \wedge g_{2p} \wedge \operatorname*{\bigsymb{\bigvee}}_{y \in Y}\mpar{ \alpha_1 = \alpha_{2p} \wedge \alpha_{2p}=\alpha_{3py} \wedge g_{3py} }
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\operatorname*{\bigsymb{\bigvee}}_{y \in Y} 
\mpar{\everymath{\displaystyle}\begin{array}{l}		
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_{2p} \land \bigwedge_\subbox{j \in J'_{2p}} \beta_{2pj} = \beta_{1xj} \land  \bigwedge_\subbox{j \in J'_{3py} \cap H} \beta_{2pj}=\beta_{3pyj} \\ g_{1 x}  \land
 R_{12}\mpar{s'_{1 x}, s'_{2p}} \wedge R_{23}\mpar{s'_{2p}, s'_{3y}} \psubst{\psi_{1 x} \uplus \psi_{2p}\uplus \psi_{3py}}
 \end{array}}
\end{multline*}

The formula is valid for all  $s_{2p} \in (s_{2p})^{p\in P}$. To build $R_{13}$ we need to rely on the disjunction of all possible path to relate $s_1$ and $s_3$:
\begin{multline*}
\operatorname*{\bigsymb{\bigvee}}_{p \in P} \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge  R_{23}\mpar{s_{2p}, s_{3}} \wedge	g_1 \wedge g_{2p} \wedge \operatorname*{\bigsymb{\bigvee}}_{y \in Y}\mpar{ \alpha_1 = \alpha_{3py} \wedge g_{3py}}}
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\operatorname*{\bigsymb{\bigvee}}_{p \in P}\operatorname*{\bigsymb{\bigvee}}_{y \in Y} 
\mpar{\everymath{\displaystyle}\begin{array}{l}		
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_{2p} \land \bigwedge_\subbox{j \in J'_{2p}} \beta_{2pj} = \beta_{1xj} \land  \bigwedge_\subbox{j \in J'_{3py} \cap H} \beta_{2pj}=\beta_{3pyj} \\ g_{1 x}  \land
 R_{12}\mpar{s'_{1 x}, s'_{2p}} \wedge R_{23}\mpar{s'_{2p}, s'_{3y}} \psubst{\psi_{1 x} \uplus \psi_{2p}\uplus \psi_{3py}}
 \end{array}}
\end{multline*}

--------------------------------------

Firstly, according to the refinement relation between  $A_1$ and $A_2$ we have:
\begin{multline*}
\bigforall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_2} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R_{12}\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}
\end{multline*}
Yet according to the refinement  relation between   $A_2$ and $A_3$ and Definition \ref{Def:OA-Refinement} we have:

\begin{multline*}
\bigforall \mpar{s_2, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_2, t_3} = \mpar{\OTx{2}{}{2}{2}, \OTx{3}{}{3}{3}} \in \fOT{s_2} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2x} = \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X},
V_2 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{2x}} \uplus \fvars{t_3} \uplus \fvars{t_2} \vdash \hspace{6cm}\\  R_{23}\mpar{s_2, s_3} \wedge 	g_2 \wedge g_3 \wedge \alpha_2 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2j} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_2=\alpha_{2 x}  \land J'_{2x}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2xj}\land g_{2 x}\land
 R_{23}\mpar{s'_{2 x}, s'_3}\psubst{\psi_{2 x} \uplus \psi_3}}\\
\end{multline*}

Note that if  either $R_{12}\mpar{s_1, s_2}$ or $R_{23}\mpar{s_2, s_3}$ is false the statement trivially holds.
So for each $s_2 \in S_2$, we remark that  $s_2$  allows to relate $s_1$ and $s_3$ if and only if $R_{12}\mpar{s_1, s_2}\land R_{23}\mpar{s_2, s_3} $. We pick a state $s_2$ relating both,  we can then the following be inferred from the above formulas:
$\forall \mpar{s_1,s_3} \in S_1 \times S_3$ and\\ $\bigsymb{\forall} \mpar{t_1, t_{3}}\!=\! \mpar{\OTx{1}{}{1}{1},\OTx{3}{}{3}{3}}\! \in\! \fOT{s_1} \times \fOT{s_3}$  
\begin{multline*}
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},\\
 V_1 \uplus V_2 \uplus V_3 \uplus \fvars{t_1} \uplus \fvars{t_2} \uplus \fvars{t_3} \uplus \biguplus_{x\in X}\fvars{t_{1x}}     \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \wedge 	g_1 \wedge g_2 \wedge   g_3\wedge  \alpha_1 = \alpha_2  \wedge \alpha_2 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2j} = \beta_{3j}} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R_{12}\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}} \wedge\\ R_{23}\mpar{s'_{2}, s'_3}\psubst{\psi_{2} \uplus \psi_3}\hspace{6cm}\\
\end{multline*}
By rearranging, simplifying and removing the variables of $A_2$ that have not effect on   $A_1$ et  $A_3$, we get the desired result.

>>>>

Pictorially, the theorem has the following form:

@@

The bisimulation relation that we consider is the one that makes $A_1$ and $A_2$ bisimilar, complemented with identity of configurations for $A_3$.
The proof relies on the fact that, by construction, all transitions of the composed automaton $A_1\subst{A_3}{k}$ are specified by open transitions of $A_1$. For the transitions that do not involve hole $k$, the transition of $A_1\subst{A_3}{k}$ is the same and simulation between $A_1$ and $A_2$ allows us to conclude directly. If the hole $k$ is involved the considered relation  implies that valuations in $A_3$ are equal (i.e., the value for each variable are the same in both valuations), after a transition we should obtain ``equal'' valuations because post-conditions are deterministic.
\begin{example} The composition of the refined \texttt{database} and the \texttt{counter} is again an open automata, depicted in Fig.~\ref{Fig:ComposeRefine}. As stated by the theorem,  it can easily be verified that this automaton is a refinement of the automaton of the figure Fig.~\ref{Def:CompOA}, i.e. the automaton resulting from the composition of the abstract \texttt{database}.
\end{example}

@@

@@

Previous works on open automata focused on equivalence relations compatible with composition.
In  \cite{10.1145/3372884.3373161}, a computable bisimulation is introduced, while in \cite{fhbisim} is introduced a weak version of the bisimulation on open automata. 
In this paper we tackle the refinement relation in the form of simulation, as is the case for the corresponding relations on labelled transition systems \cite{10.1007/3-540-46428-X_19}. But unlike the standard simulation we deal with symbolic and open models. In \cite{Zhang2014}, the authors exploit transition systems to reason about the systems  that are partially specified by using variables, making  the state space potentially infinite.

The simulation  ensures the preservation of   
safety properties as deadlock-freeness  and, more generally, all 
all linear temporal logic properties.  The issue is about the properties of simulation, in particular closure properties such as preservation under composition. Some works   have addressed the problem of non-introduction of blocking by composition. For example, in \cite{DIHEGO2020110598} it is presented a refinement relation on process algebra based on failure semantics that is basically different from the (bi)simulation\cite{10.5555/640428.640430}.

@@

\newpage
\TODO{[RABEA] proof}
Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. According to the refinement relation between  $A_1$ and $A_2$, we have  the following formula that we denote ($*$):\\
for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}	
On the other hand, according to the refinement relation between   $A_2$ and $A_3$, we know condition 3 is satisfied, $\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3$ and $\bigsymb{\forall} \mpar{t_{2p}, t_3} \in \fOT{s_{2p}} \times \fOT{s_3}$ and in particular for  the set $(t_{2px})^{x\in X_p}$ so we  have the following that we denote ($**$):
\begin{multline*}
\bigsymb{\forall} \mpar{t_{2px}, t_3} = \mpar{\OTx{2px}{}{2px}{2px}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2pxy} = \OTx{2px}{y}{2pxy}{2pxy} \in \fOT{s_{2p}}}^{y \in Y_X},
V_2 \uplus V_3\uplus \biguplus_{y\in Y}\fvars{t_{2pxy}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{y\in Y_X}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}\\
\end{multline*}
Again according to the refinement  relation between  $A_1$ and $A_2$ and condition 3  the following is satisfied $\bigforall \mpar{s_1, s_{2p}} \in S_1 \times S_2$ , in particular over the family of transitions in $(t_{2pxy})^{y\in Y_X}$:
\begin{multline*}
\bigsymb{\forall} \mpar{t_1, t_{2pxy}} = \mpar{\OTx{1}{}{1}{1}, \OTx{2pxy}{}{2pxy}{2pxy}} \in \fOT{s_1} \times \fOT{s_{2p}} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1z} = \OTx{1}{z}{1z}{1z} \in \fOT{s_1}}^{z \in Z},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pxyj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}
\end{multline*}

This is verified for all $y \in  Y_X$.
\begin{multline*}
\bigsymb{\bigwedge_{y \in  Y_X}}
\Bigg[R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy}
\implies \hspace{6cm}\\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}\Bigg]
\end{multline*}

The conjunction of formulas ($*$) and ($**$) gives:
\begin{multline*}
\bigsymb{\bigwedge_{y \in  Y_X}}
\Bigg[
\mpar{R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy}} \wedge \mpar{ R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 } 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}} \wedge \\
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}
 \Bigg]
\end{multline*}
By rearranging the formula we get:
\begin{multline*}
\bigsymb{\bigwedge_{y \in  Y_X}}
\Bigg[
R_{12}\mpar{s_1, s_{2p}} \wedge  R_{23}\mpar{s_{2p}, s_3} \wedge g_1 \wedge g_3 \wedge 	g_{2px}\wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} \wedge     \alpha_{2px}= \alpha_3 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}} \wedge \\
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}  \psubst{\psi_{2p xy} \uplus \psi_3}}
 \Bigg]
\end{multline*}
\TODO{ Je suis bloque a ce niveau: dans la conclusion on a $\alpha_{2px}=\alpha_{2p xy}$ qui est dans les hypotheses.  
}

The formula is valid for all  $s_{2p} \in (s_{2p})^{p\in P}$. To build $R_{13}$ we need to rely on the disjunction of all possible path to relate $s_1$ and $s_3$:

Note that if  either $R_{12}\mpar{s_1, s_2}$ or $R_{23}\mpar{s_2, s_3}$ is false the statement trivially holds.
So for each $s_2 \in S_2$, we remark that  $s_2$  allows to relate $s_1$ and $s_3$ if and only if $R_{12}\mpar{s_1, s_2}\land R_{23}\mpar{s_2, s_3} $. We pick a state $s_2$ relating both,  we can then the following be inferred from the above formulas:
$\forall \mpar{s_1,s_3} \in S_1 \times S_3$ and\\ $\bigsymb{\forall} \mpar{t_1, t_{3}}\!=\! \mpar{\OTx{1}{}{1}{1},\OTx{3}{}{3}{3}}\! \in\! \fOT{s_1} \times \fOT{s_3}$

By rearranging, simplifying and removing the variables of $A_2$ that have not effect on   $A_1$ et  $A_3$, we get the desired result.
